---
title: "Exploratory Bar Graphs - JHarte Client"
author: "Sam Pritchard"
date: "December 20, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(dplyr)
require(lubridate)
require(knitr)
```



```{r,echo=FALSE}

# NEED TO MAKE IT YTD VS. YTD!!!!

#data17 <- read.csv("data/Detailed Sales Report 17 full.csv", stringsAsFactors = FALSE)
#data16 <- read.csv("data/Detailed Sales Report 16.csv", stringsAsFactors = FALSE)
#data16 <- data16[,-c(18, 19, 21)]

#data.clean <- rbind(data17, data16)
#data.clean$Total.Sales <- as.numeric(data.clean$Total.Sales)
#data.clean <- filter(data.clean, !is.na(Total.Sales)) #removes the sums that had $ signs in them
#print(paste("sum of Total.Sales is", sum(data.clean$Total.Sales)))
#data.clean$Date.Sold <- mdy_hm(data.clean$Date.Sold)
#data.clean$Year <- year(data.clean$Date.Sold)
#data.clean$Month <- month(data.clean$Date.Sold)
#data.clean$Department <- trimws(data.clean$Department)
#data.clean$Category <- trimws(data.clean$Category)
#data.clean$Supplier <- trimws(data.clean$Supplier)
#data.clean$Item <- trimws(data.clean$Item)
#print(head(data.clean))
#data.clean$Dept.By.Year <- paste(data.clean$Department, data.clean$Year)
#data.clean$Categ.By.Year <- paste(data.clean$Category, data.clean$Year)

import.jh.sql.connect <- function(db.name.in.sql.server){
  library(odbc)
  library(DBI)
  
  con <- DBI::dbConnect(odbc::odbc(),
                        Driver = "SQL Server",
                        Server = "LAPTOP-JHRVTF04\\SQLEXPRESS",
                        Database = db.name.in.sql.server,
                        Trusted_Connection = "True")
  con
}

import.jh.detailed.sales <- function(){
  joined.df <- DBI::dbGetQuery(con,'
                                SELECT TransactionEntry.ItemID, TransactionEntry.Price, 
                                 TransactionEntry.TransactionTime, TransactionEntry.Quantity, 
                                  TransactionEntry.TransactionNumber, Item.Description, 
                                 Department.Name AS "Department", Category.Name AS "Category", 
                                 SupplierName AS "Supplier", Cashier.Name AS "CashierName", 
                                 Register.Number AS "Register", Register.Description AS "RegLocation"

                                 FROM TransactionEntry
                                 
                                 JOIN Item ON TransactionEntry.ItemID = Item.ID
                                 JOIN Department ON Item.DepartmentID = Department.ID
                                 JOIN Category ON Item.CategoryID = Category.ID
                                 JOIN Supplier ON Item.SupplierID = Supplier.ID
                                 JOIN [Transaction] ON TransactionEntry.TransactionNumber = [Transaction].TransactionNumber
                                 JOIN Batch ON [Transaction].BatchNumber = Batch.BatchNumber
                                 JOIN Register ON Batch.RegisterID = Register.ID
                                 JOIN Cashier ON [Transaction].CashierID = Cashier.ID')
  joined.df
  }

clean.jh.detailed.sales <- function(sql.df){
  library(dplyr)
  library(lubridate)
    
    sql.df <- sql.df %>%
      rename(Sold.Price = Price, Cashier = CashierName, Date.Sold = TransactionTime,
             Qty.Sold = Quantity, Transaction = TransactionNumber) %>%
      mutate(Total.Sales = Qty.Sold*Sold.Price) %>%
      select(Department, Category, Supplier, ItemID, Description, Qty.Sold, Sold.Price, 
             Total.Sales, Transaction, Date.Sold, Register, Cashier)
    
    sql.df$Date.Sold <- ymd_hms(sql.df$Date.Sold)
    sql.df$Year <- year(sql.df$Date.Sold)
    sql.df$Month <- month(sql.df$Date.Sold)
    sql.df$Department <- trimws(sql.df$Department)
    sql.df$Category <- trimws(sql.df$Category)
    sql.df$Supplier <- trimws(sql.df$Supplier)
    sql.df$Item <- trimws(sql.df$Item)
    #print(head(sql.df))
    sql.df$Dept.By.Year <- paste(sql.df$Department, sql.df$Year)
    sql.df$Categ.By.Year <- paste(sql.df$Category, sql.df$Year)
    sql.df
}

con <- import.jh.sql.connect("Sam's Store")

new.df <- import.jh.detailed.sales() # uses con but not as argument

data.clean <- clean.jh.detailed.sales(new.df)


dept.agg <- aggregate(Total.Sales ~ Dept.By.Year + Department + Categ.By.Year + Category + Year, data.clean, sum)
```

#SUPPLIERS
Of 435 Suppliers the top 10% (44 of them) make up 55.1% of sales
The top 20% (88 of them) make up 74.6% of sales
30% makes up 85.3%
40% makes up 92.2%
50% makes up 96.4%

In $ the bottom 25% brought in $10,671 in 2017, which is 109 suppliers
```{r, echo=FALSE}
data.clean.supplier <- data.clean
data.clean.supplier$Supp.By.Year <- paste(data.clean.supplier$Supplier, data.clean.supplier$Year)
supplier.agg <- aggregate(Total.Sales ~ Supp.By.Year + Supplier + Year, data.clean.supplier, sum)
supplier.agg17 <- filter(supplier.agg, Year == 2017)
supp17sum <- sum(supplier.agg17$Total.Sales)
supplier.agg17 <- mutate(supplier.agg17, "Perc.Whole" = round((Total.Sales/supp17sum)*100, 1))
supplier.agg17 <- arrange(supplier.agg17, desc(Total.Sales))
print(head(supplier.agg17))
print(nrow(supplier.agg17))
kable(supplier.agg17)

sum(supplier.agg17$Perc.Whole[1:44]) #10%

sum(supplier.agg17$Perc.Whole[1:88]) #20%
sum(supplier.agg17$Perc.Whole[1:132]) #30%
sum(supplier.agg17$Perc.Whole[1:176]) #40%
sum(supplier.agg17$Perc.Whole[1:218]) #50%

sum(supplier.agg17$Perc.Whole[(435-44):435]) #bottom 10%
sum(supplier.agg17$Perc.Whole[326:435]) #bottom 25%

supplier.agg16 <- filter(supplier.agg, Year == 2016)
sup.merge16 <- supplier.agg16[,c(2,4)]
colnames(sup.merge16) <- c("Supplier", "2016")
sup.merge17 <- supplier.agg17[,c(2,4)]
colnames(sup.merge17) <- c("Supplier", "2017")
sup.merge <- merge(sup.merge17, sup.merge16)
sup.merge2 <- mutate(sup.merge, "Growth" = `2017` - `2016`, "Perc.Growth" = paste0(round((Growth/`2016`)*100,1),"%"))
supplier.agg17 <- arrange(supplier.agg17, desc(Total.Sales))
b25suppliers17 <- supplier.agg17[326:435,]
b25growth <- filter(sup.merge2, Supplier %in% b25suppliers17$Supplier)

b25growth <- arrange(b25growth, desc(`2017`))
library(knitr)
kable(b25growth)



```

## Bar Graphs Departments

```{r, echo=FALSE}


dept.bar <- ggplot(data = dept.agg, aes(Dept.By.Year)) +
  geom_bar(aes(fill=Department, weight=Total.Sales)) + 
  coord_flip() + theme(legend.position = "none")

print(dept.bar)


```

## Remove Very Small Depts

Removed the following categories: "POSTAGE", "MISCELLANEOUS PAPER", "MISCELLANEOUS", "GIFT CARD", "DISPLAY"

```{r, echo=FALSE}
small.depts <- c("POSTAGE", "MISCELLANEOUS PAPER", "MISCELLANEOUS", "GIFT CARD", "DISPLAY", "")

dept.not.small.agg <- dept.agg[!(dept.agg$Department %in% small.depts),]

dept.not.small.bar <- ggplot(data = dept.not.small.agg, aes(Dept.By.Year)) +
  geom_bar(aes(fill=Department, weight=Total.Sales)) + 
  coord_flip() + theme(legend.position = "none")

print(dept.not.small.bar)
```

### Jewelry by Category

Slight declines in most major categories but the one in bracelets is larger.  The declines follow the trend for the category

```{r, echo=FALSE}

#jewelry.agg <- filter(dept.not.small.agg, Department == "JEWELRY")
cat.bar <- ggplot(data = dept.agg, aes(Categ.By.Year)) +
  geom_bar(aes(fill=Category, weight=Total.Sales)) + 
  coord_flip() + theme(legend.position = "none")

print(cat.bar)


```

### General Merchandise by Category
The Categories HOUSEWARES, CRAFTS, and ART GLASS were the highest for General Merchandise.  Housewares also grew substantially over last year, as did PERS ACCES/BAGS/POUCHES.

As in some other categories, fewer items are uncategorized which may make growth look better than in really is in 2017.
Best categories in ters

```{r, echo=FALSE}

gmerch.agg <- filter(dept.not.small.agg, Department == "GENERAL MERCHANDISE")
gmerch.bar <- ggplot(data = gmerch.agg, aes(Categ.By.Year)) +
  geom_bar(aes(fill=Category, weight=Total.Sales)) + 
  coord_flip() + theme(legend.position = "none")

print(gmerch.bar)


```

### Counter Card by Category

BIRTHDAY is the largest category by far but saw a slight decline.

The second highest is BLANK cards which also saw an increase in 2017

```{r, echo=FALSE, fig.height=11}

counterc.agg <- filter(dept.not.small.agg, Department == "COUNTER CARD")
counterc.bar <- ggplot(data = counterc.agg, aes(Categ.By.Year)) +
  geom_bar(aes(fill=Category, weight=Total.Sales)) + 
  coord_flip() + theme(legend.position = "none")

print(counterc.bar)


```

### Toys by Category

IMAGINATIVE PLAY was the top toy category last year and it has continued to an even bigger year as top seller again in 2017. 

PLUSH similarly was a really big seller last year (#2) and saw growth on top of that. 

Other Categories with big growth were CREATIVE LEARNING, BABY, OUTDOOR FUN.

Notable declines occurred in PUZZLES, PUPPETS, and FINGER PUPPETS

Note that there were more uncategorized TOYS purchases in 2016, so that likely inflates some of the 2017 categories 

```{r, echo=FALSE}

toys.agg <- filter(dept.not.small.agg, Department == "TOYS")
toys.bar <- ggplot(data = toys.agg, aes(Categ.By.Year)) +
  geom_bar(aes(fill=Category, weight=Total.Sales)) + 
  coord_flip() + theme(legend.position = "none")

print(toys.bar)


```


### Books by Category

TOY BOOK HARDBACK is the highest in the category and brings in roughly $80K per year

```{r, echo=FALSE}

books.agg <- filter(dept.not.small.agg, Department == "BOOKS")
books.bar <- ggplot(data = books.agg, aes(Categ.By.Year)) +
  geom_bar(aes(fill=Category, weight=Total.Sales)) + 
  coord_flip() + theme(legend.position = "none")

print(books.bar)


```
