---
title: "Supplier Snapshot - Whale's Tale"
author: "JHarte & Associates + Allocate Analytics"
date: "January 15, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(ggplot2)
require(lubridate)
require(knitr)
require(fpp2)

#contents 2019

#1 Autoplot of data for selected time period to show time period and a great viz out of the gate
#2 text info about the products and the suppliers broken down

#3 Text about the 80-20 rule - which suppliers' products generating the most sales and the least

#####
freemium.end.date <- as.Date("2018-12-31") ##### Date to end the data 
#freemium.end.date <- as.Date("2099-12-31") #### removes no dates but does remove NAs
#####

#next if-else
if(file.exists(paste0("data/sales.data.cache", Sys.Date(), ".csv"))){
  data.clean <- read.csv(paste0("data/sales.data.cache", Sys.Date(), ".csv"), stringsAsFactor = FALSE)
} else {
  
import.jh.sql.connect <- function(db.name.in.sql.server){
  library(odbc)
  library(DBI)
  
  con <- DBI::dbConnect(odbc::odbc(),
                        Driver = "SQL Server",
                        Server = "LAPTOP-JHRVTF04\\SQLEXPRESS",
                        Database = db.name.in.sql.server,
                        Trusted_Connection = "True")
  con
}
import.jh.detailed.sales <- function(){
  joined.df <- DBI::dbGetQuery(con,'
                                SELECT TransactionEntry.ItemID, TransactionEntry.Price, 
                                 TransactionEntry.TransactionTime, TransactionEntry.Quantity, 
                                  TransactionEntry.TransactionNumber, Item.Description, 
                                 Department.Name AS "Department", Category.Name AS "Category", 
                                 SupplierName AS "Supplier", Cashier.Name AS "CashierName", 
                                 Register.Number AS "Register", Register.Description AS "RegLocation"
                                 FROM TransactionEntry
                                 
                                 JOIN Item ON TransactionEntry.ItemID = Item.ID
                                 JOIN Department ON Item.DepartmentID = Department.ID
                                 JOIN Category ON Item.CategoryID = Category.ID
                                 JOIN Supplier ON Item.SupplierID = Supplier.ID
                                 JOIN [Transaction] ON TransactionEntry.TransactionNumber = [Transaction].TransactionNumber
                                 JOIN Batch ON [Transaction].BatchNumber = Batch.BatchNumber
                                 JOIN Register ON Batch.RegisterID = Register.ID
                                 JOIN Cashier ON [Transaction].CashierID = Cashier.ID')
  joined.df
  }
clean.jh.detailed.sales <- function(sql.df){
  library(dplyr)
  library(lubridate)
    
    sql.df <- sql.df %>%
      rename(Sold.Price = Price, Cashier = CashierName, Date.Sold = TransactionTime,
             Qty.Sold = Quantity, Transaction = TransactionNumber) %>%
      mutate(Total.Sales = Qty.Sold*Sold.Price) %>%
      select(Department, Category, Supplier, ItemID, Description, Qty.Sold, Sold.Price, 
             Total.Sales, Transaction, Date.Sold, Register, Cashier)
    
    sql.df$Date.Sold <- ymd_hms(sql.df$Date.Sold)
    sql.df$Year <- year(sql.df$Date.Sold)
    sql.df$Month <- month(sql.df$Date.Sold)
    sql.df$Department <- trimws(sql.df$Department)
    sql.df$Category <- trimws(sql.df$Category)
    sql.df$Supplier <- trimws(sql.df$Supplier)
    sql.df$Item <- trimws(sql.df$Item)
    #print(head(sql.df))
    sql.df$Dept.By.Year <- paste(sql.df$Department, sql.df$Year)
    sql.df$Categ.By.Year <- paste(sql.df$Category, sql.df$Year)
    sql.df
}

con <- import.jh.sql.connect("Sam's Store")
new.df <- import.jh.detailed.sales() # uses con but not as argument
data.clean <- clean.jh.detailed.sales(new.df)

write.csv(data.clean, paste0("data/sales.data.cache", Sys.Date(), ".csv"), row.names = FALSE)
}

#read in another function independent of caching
exploratory.jh.time.series <- function(clean.df, freq = 365){
  require(lubridate)
  require(dplyr)
  clean.df <- arrange(clean.df, Date.Sold)
  clean.df$Day <- date(clean.df$Date.Sold) # df should have been arranged in cleaning step
  start.year <- year(clean.df$Day)[1]
  if(freq == 365){
    start.day <- date(head(clean.df$Day,1)) - floor_date(head(clean.df$Day,1), unit = "year")+1 # subtracts earliest day from first day of year plus one
    sales.agg <- aggregate(Total.Sales ~ Day, clean.df, sum)
    all.dates <- seq.Date(from = as.Date(min(sales.agg$Day, na.rm = TRUE)), to = as.Date(max(sales.agg$Day, na.rm = TRUE)),by = 1)
    sales.agg.all <- left_join(data.frame(Day = all.dates), sales.agg)
    sales.ts <- ts(sales.agg.all$Total.Sales, start = c(start.year, start.day), frequency = 365)
  }
  if(freq == 12){
    start.month <- month(head(clean.df$Month,1))
    sales.agg <- aggregate(Total.Sales ~ Month + Year, clean.df, sum)
    #print(head(sales.agg, 25))
    sales.agg.all <- left_join(sales.agg, data.frame(Month = 1:12))
    sales.ts <- ts(sales.agg.all$Total.Sales, start = c(start.year, start.month), frequency = 12)
  }
  sales.ts[is.na(sales.ts)] <- 0
  
  sales.ts
}

data.clean <- filter(data.clean, Date.Sold <= freemium.end.date) ##### check
data.both <- data.clean
#dept.agg <- aggregate(Total.Sales ~ Dept.By.Year + Department + Categ.By.Year + Category + Year, data.both, sum)

data.both.supplier <- data.both
data.both.supplier$Supp.By.Year <- paste(data.both.supplier$Supplier, data.both.supplier$Year)
supplier.agg <- aggregate(Total.Sales ~ Supp.By.Year + Supplier + Year, data.both.supplier, sum)
cy <- year(freemium.end.date)
py <- year(freemium.end.date)-1
supplier.agg.cy <- filter(supplier.agg, Year == cy)
supp.cy.sum <- sum(supplier.agg.cy$Total.Sales)
supplier.agg.cy <- mutate(supplier.agg.cy, "Perc.Whole" = round((Total.Sales/supp.cy.sum)*100, 2))
supplier.agg.cy <- arrange(supplier.agg.cy, desc(Total.Sales))

prod.uniq <- length(unique(data.both$ItemID))
cat.uniq <- length(unique(data.both$Category))
dept.uniq <- length(unique(data.both$Department))
```


###This report enables a business to look at their suppliers a couple different ways.  Some of the questions it can help answer are the following:

*	Are there suppliers generating so little in sales they’re not worth the shelf space, inventory, and communication I’m giving them?

*	If I were to get rid of my smallest 50% of suppliers, how much would that represent in sales?

*	Which suppliers brought in less this year than the year before?

*	Which products are selling best for my best suppliers?

*	Are there certain times of year when a top supplier’s products are selling better?

Note: this report relies on data from `r as.Date(min(data.clean$Date.Sold, na.rm = TRUE))` to `r as.Date(max(data.clean$Date.Sold, na.rm = TRUE))`. 



```{r, allin, echo = FALSE, warning=FALSE, message=FALSE, fig.height=3}
allin.agg <- aggregate(Total.Sales ~ Year, supplier.agg, sum)

#####just a test
#allin.agg <- filter(allin.agg, Year >= 2018)

#####
years.incl <- unique(allin.agg$Year)  # important because picking color palettes manually for lines

all.ts <- exploratory.jh.time.series(data.clean, freq = 12)

if(length(years.incl) == 1){

  man.pal <- c("#67a9cf") #manual palette
  
  all.bar <- ggplot(data = allin.agg, aes(y = Total.Sales, x = factor(Year), fill = factor(Year))) + 
    geom_bar(stat = "identity") + scale_y_continuous(labels = scales::dollar) + 
    scale_fill_manual(name = "Year", values = man.pal) + theme_dark()+
    labs(x = "Year", y = "Sales", title = "Total Sales from All Suppliers by Year") + 
    theme(legend.position="none")

year1.ts <- all.ts

all.line <- autoplot(all.ts) + autolayer(year1.ts, color = man.pal[1], size = 3) + theme_dark()+ 
  labs(y = "Sales", title = "Total Sales from All Suppliers by Month") +
  scale_y_continuous(labels = scales::dollar)

}

if(length(years.incl) == 2){

  man.pal <- c("#f7f7f7", "#67a9cf") #manual palette
  
  all.bar <- ggplot(data = allin.agg, aes(y = Total.Sales, x = factor(Year), fill = factor(Year))) + 
  geom_bar(stat = "identity") + scale_y_continuous(labels = scales::dollar) + 
  scale_fill_manual(name = "Year", values = man.pal) + theme_dark()+
  labs(x = "Year", y = "Sales", title = "Total Sales from All Suppliers by Year") + theme(legend.position="none")


year1.ts <- window(all.ts, end = c(years.incl[2],1))
year2.ts <- window(all.ts, start = c(years.incl[2],1))
all.line <- autoplot(all.ts) + autolayer(year1.ts, color = man.pal[1], size = 3) + 
  autolayer(year2.ts, color = man.pal[2], size = 3) + theme_dark()+ 
  labs(y = "Sales", title = "Total Sales from All Suppliers by Month") +
  scale_y_continuous(labels = scales::dollar)
}

if(length(years.incl) == 3){

all.bar <- ggplot(data = allin.agg, aes(y = Total.Sales, x = factor(Year), fill = factor(Year))) + 
  geom_bar(stat = "identity") + scale_y_continuous(labels = scales::dollar) + 
  scale_fill_brewer(name = "Year", palette = "RdBu") + theme_dark()+
  labs(x = "Year", y = "Sales", title = "Total Sales from All Suppliers by Year") + theme(legend.position="none")

man.pal <- c("#ef8a62", "#f7f7f7", "#67a9cf") #manual palette
year1.ts <- window(all.ts, end = c(years.incl[2],1))
year2.ts <- window(all.ts, start = c(years.incl[2],1), end = c(years.incl[3],1))
year3.ts <- window(all.ts, start = c(years.incl[3],1))

all.line <- autoplot(all.ts) + autolayer(year1.ts, color = man.pal[1], size = 3) + 
  autolayer(year2.ts, color = man.pal[2], size = 3) + 
  autolayer(year3.ts, color = man.pal[3], size = 3) + theme_dark() + 
  labs(y = "Sales", title = "Total Sales from All Suppliers by Month") +
  scale_y_continuous(labels = scales::dollar)

}

if(length(years.incl) == 4){
#c(#ca0020, #f4a582, #92c5de, #0571b0)
  
  all.bar <- ggplot(data = allin.agg, aes(y = Total.Sales, x = factor(Year), fill = factor(Year))) + 
  geom_bar(stat = "identity") + scale_y_continuous(labels = scales::dollar) + 
  scale_fill_brewer(name = "Year", palette = "RdBu") + theme_dark()+
  labs(x = "Year", y = "Sales", title = "Total Sales from All Suppliers by Year") + theme(legend.position="none")

man.pal <- c("#ca0020", "#f4a582", "#92c5de", "#0571b0") #manual palette
year1.ts <- window(all.ts, end = c(years.incl[2],1))
year2.ts <- window(all.ts, start = c(years.incl[2],1), end = c(years.incl[3],1))
year3.ts <- window(all.ts, start = c(years.incl[3],1), end = c(years.incl[4],1))
year4.ts <- window(all.ts, start = c(years.incl[4],1))

all.line <- autoplot(all.ts) + autolayer(year1.ts, color = man.pal[1], size = 3) + 
  autolayer(year2.ts, color = man.pal[2], size = 3) + 
  autolayer(year3.ts, color = man.pal[3], size = 3) + 
  autolayer(year4.ts, color = man.pal[4], size = 3) + theme_dark()+ 
  labs(y = "Sales", title = "Total Sales from All Suppliers by Month") +
  scale_y_continuous(labels = scales::dollar)
}

if(length(years.incl) == 5){
#c(#ca0020, #f4a582, #f7f7f7, #92c5de, #0571b0)
  
  all.bar <- ggplot(data = allin.agg, aes(y = Total.Sales, x = factor(Year), fill = factor(Year))) + 
  geom_bar(stat = "identity") + scale_y_continuous(labels = scales::dollar) + 
  scale_fill_brewer(name = "Year", palette = "RdBu") + theme_dark()+
  labs(x = "Year", y = "Sales", title = "Total Sales from All Suppliers by Year") + theme(legend.position="none")

man.pal <- c("#ca0020", "#f4a582", "#f7f7f7","#92c5de", "#0571b0") #manual palette
year1.ts <- window(all.ts, end = c(years.incl[2],1))
year2.ts <- window(all.ts, start = c(years.incl[2],1), end = c(years.incl[3],1))
year3.ts <- window(all.ts, start = c(years.incl[3],1), end = c(years.incl[4],1))
year4.ts <- window(all.ts, start = c(years.incl[4],1), end = c(years.incl[5],1))
year5.ts <- window(all.ts, start = c(years.incl[5],1))

all.line <- autoplot(all.ts) + autolayer(year1.ts, color = man.pal[1], size = 3) + 
  autolayer(year2.ts, color = man.pal[2], size = 3) + 
  autolayer(year3.ts, color = man.pal[3], size = 3) + 
  autolayer(year4.ts, color = man.pal[4], size = 3) +
  autolayer(year5.ts, color = man.pal[5], size = 3) + theme_dark()+ 
  labs(y = "Sales", title = "Total Sales from All Suppliers by Month") +
  scale_y_continuous(labels = scales::dollar)
}

#autoplot(all.ts) + theme_dark() + geom_line(aes(color = Year), size = 4) + scale_color_brewer(palette = "RdBu")



```

## Summary
**For the selected time period there were `r prod.uniq` distinct products (SKUs) sold in `r cat.uniq` categories across `r dept.uniq` departments from `r nrow(supplier.agg.cy)` different suppliers.**

## Overview of Sales by Year and by Month for All Suppliers

```{r, echo=FALSE}
print(all.bar)
print(all.line)

```

## Chart of Sales by LARGEST X% of Suppliers for `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`

Some people are familiar with the 80-20 rule, or the Pareto Principle.  It states that in many areas of life and business 80% of the results come from 20% of the possible sources.  Examples include 80% of complaints come from 20% of customers or 80% of people in a country might only have 20% of the assets.  In our experience in business, the ratios aren’t quite that precise but it is often true that a large portion of sales can often come from a small portion of suppliers.  These next charts are meant to identify and visualize those portions for this particular business.

```{r, echo=FALSE}

#create data frame with columns 
      #1) Top Percentage of all suppliers in increments of 1% 
      #2) Total sales at each increment 
      #3) Total units (rows in detailed sales report) at each increment 
      #4) Number of items associated with those suppliers (at least those sold)
# begin with data by suppliers showing Total.Sales, Items, units sold
increment.rows <- data.frame()
perc.inc.df <- for(h in 1:100){
  one.to.h.sum <- sum(supplier.agg.cy$Total.Sales[1:((h/100)*nrow(supplier.agg.cy))])
  row.h <- c(h, round(nrow(supplier.agg.cy)*(h/100)), one.to.h.sum)
  #print(row.h)
  increment.rows <- rbind(increment.rows, row.h)
}

colnames(increment.rows) <- c("Perc. of LARGEST Suppliers", "Number of Suppliers", "Sales")
increment.rows <- mutate(increment.rows, "Perc. of Sales" = (increment.rows[,3]/tail(increment.rows[,3],1))*100)

#decile.rows
decile.rows <- data.frame()
h.by.tens <- seq(from = 10, to = 100, by = 10)
#print(h.by.tens)
perc.inc.df <- for(h in h.by.tens){
  one.to.h.sum <- sum(supplier.agg.cy$Total.Sales[1:((h/100)*nrow(supplier.agg.cy))])
  row.h <- c(h, round(nrow(supplier.agg.cy)*(h/100)), one.to.h.sum)
  #print(row.h)
  decile.rows <- rbind(decile.rows, row.h)
}

colnames(decile.rows) <- c("Perc. of LARGEST Suppliers", "Number of Suppliers", "Sales")
decile.rows <- mutate(decile.rows, "Perc. of Sales" = (decile.rows[,3]/tail(decile.rows[,3],1))*100)
#supp.conc <- ggplot(increment.rows, aes(x=`Perc. of LARGEST Suppliers`, y=Sales, group= 1)) +
#  geom_line(size=2, color="blue")

supp.conc.bars <- ggplot(decile.rows, aes(x=`Perc. of LARGEST Suppliers`, y=Sales, group= 1, fill = Sales)) +
  geom_bar(stat = "identity") + theme_dark() +
  scale_fill_distiller(palette = "Blues", direction = 1) +
  scale_y_continuous(labels = scales::dollar) + 
  scale_x_continuous(breaks = seq(from = 10, to = 100, by = 10), labels = paste0(h.by.tens, "%")) +
  guides(fill = FALSE)

###print(supp.conc.bars)
#pretty up decile rows data for use in kable
decile.rows$Sales <- paste0("$",prettyNum(round(decile.rows$Sales), big.mark = ","))
decile.rows$`Perc. of LARGEST Suppliers` <- paste0(decile.rows$`Perc. of LARGEST Suppliers`, "%")
decile.rows$`Perc. of Sales` <- paste0(round(decile.rows$`Perc. of Sales`, 1), "%")
decile.rows$`Number of Suppliers` <- as.character(decile.rows$`Number of Suppliers`)
inline1 <- decile.rows[1,1]
inline2 <- decile.rows[1,2]
inline3 <- decile.rows[1,3]
inline4 <- decile.rows[1,4]

#inline30.1 <- decile.rows[3,1]
inline30.3 <- decile.rows[3,3]
inline30.4 <- decile.rows[3,4]
```

This chart can be read as "The top 30% of suppliers measured by dollars brought in `r inline30.3`, which is `r inline30.4` of all sales in the most recent year included in the data."

```{r, echo=FALSE}
print(supp.conc.bars)

```

## Table of Sales by LARGEST X% of Suppliers for `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`

This table can be read as "the top `r inline1` percent of suppliers measured by dollars, which is `r inline2` supplier(s), brought in `r inline3` in sales, which is `r inline4` of all sales for the business in the most recent year included in the data."

```{r, echo=FALSE}
kable(decile.rows)
#increment.rows$Sales <- paste0("$",prettyNum(round(increment.rows$Sales), big.mark = ","))
#increment.rows$`Perc. of LARGEST Suppliers` <- paste0(increment.rows$`Perc. of LARGEST Suppliers`, "%")
#increment.rows$`Perc. of Sales` <- paste0(round(increment.rows$`Perc. of Sales`, 1), "%")
#increment.rows$`Number of Suppliers` <- as.character(increment.rows$`Number of Suppliers`)
#kable(increment.rows)
```





```{r, echo=FALSE}
### code for building data by 1% rather than by 10% ###
#supplier.agg.cy.bottom <- arrange(supplier.agg.cy, Total.Sales)
#increment.rows.b <- data.frame()
#perc.inc.df <- for(h in 1:100){
#  one.to.h.sum <- sum(supplier.agg.cy.bottom$Total.Sales[1:((h/100)*nrow(supplier.agg.cy.bottom))])
#  row.h <- c(h, round(nrow(supplier.agg.cy.bottom)*(h/100)), one.to.h.sum)
  #print(row.h)
#  increment.rows.b <- rbind(increment.rows.b, row.h)
#}
#colnames(increment.rows.b) <- c("Perc. of SMALLEST Suppliers", "Number of Suppliers", "Sales")
#increment.rows.b <- mutate(increment.rows.b, "Perc. of Sales" = (increment.rows.b[,3]/tail(increment.rows.b[,3],1))*100)

#supp.conc.b <- ggplot(increment.rows.b, aes(x=`Perc. of SMALLEST Suppliers`, y=Sales, group= 1)) +
#  geom_line(size=2, color="blue")
#print(supp.conc.b)

###create decile.rows.b ###
supplier.agg.cy.b <- arrange(supplier.agg.cy, Total.Sales) #b is for bottom, these are the lowest performing ones
decile.rows.b <- data.frame()
h.by.tens <- seq(from = 10, to = 100, by = 10)
#print(h.by.tens)
perc.inc.df <- for(h in h.by.tens){
  one.to.h.sum <- sum(supplier.agg.cy.b$Total.Sales[1:((h/100)*nrow(supplier.agg.cy.b))])
  row.h <- c(h, round(nrow(supplier.agg.cy.b)*(h/100)), one.to.h.sum)
  #print(row.h)
  decile.rows.b <- rbind(decile.rows.b, row.h)
}

colnames(decile.rows.b) <- c("Perc. of LARGEST Suppliers", "Number of Suppliers", "Sales")
decile.rows.b <- mutate(decile.rows.b, "Perc. of Sales" = (decile.rows.b[,3]/tail(decile.rows.b[,3],1))*100)

#new one with bars
supp.conc.bars.b <- ggplot(decile.rows.b, aes(x=`Perc. of LARGEST Suppliers`, y=Sales, group= 1, fill = Sales)) +
  geom_bar(stat = "identity") + theme_dark() +
  scale_fill_distiller(palette = "Blues", direction = 1) +
  scale_y_continuous(labels = scales::dollar) + 
  scale_x_continuous(breaks = seq(from = 10, to = 100, by = 10), labels = paste0(h.by.tens, "%")) +
  guides(fill = FALSE) + labs(x = "Perc. of SMALLEST Suppliers")

decile.rows.b$Sales <- paste0("$",prettyNum(round(decile.rows.b$Sales), big.mark = ","))
decile.rows.b$`Perc. of LARGEST Suppliers` <- paste0(decile.rows.b$`Perc. of LARGEST Suppliers`, "%")
decile.rows.b$`Perc. of Sales` <- paste0(round(decile.rows.b$`Perc. of Sales`, 1), "%")
decile.rows.b$`Number of Suppliers` <- as.character(decile.rows.b$`Number of Suppliers`)
inline50.1 <- decile.rows.b[5,1]
inline50.2 <- decile.rows.b[5,2]
inline50.3 <- decile.rows.b[5,3]
inline50.4 <- decile.rows.b[5,4]
```

## Chart of Sales by SMALLEST X% of Suppliers for `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`

The previous charts showed how much the largest suppliers generated.  These show how little the smallest have generated.  It shows the same information but in the opposite direction. It can be read starting at the x-axis on the bottom "The smallest `r inline50.1` of all suppliers accounted for `r inline50.3` in sales"

```{r, echo=FALSE}

print(supp.conc.bars.b)

```

## Table of Sales by SMALLEST X% of Suppliers for `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`

This table contains the same data as the chart above in a different format. It can be read starting at the axis on the bottom "The smallest `r inline50.1` of all suppliers, which is `r inline50.2` suppliers, accounted for `r inline50.3` in sales, which is `r inline50.4` of the whole business."

```{r, echo=FALSE}
kable(decile.rows.b)
#increment.rows.b$Sales <- paste0("$",prettyNum(round(increment.rows.b$Sales), big.mark = ","))
#increment.rows.b$`Perc. of SMALLEST Suppliers` <- paste0(increment.rows.b$`Perc. of SMALLEST Suppliers`, "%")
#increment.rows.b$`Perc. of Sales` <- paste0(round(increment.rows.b$`Perc. of Sales`, 2), "%")
#increment.rows.b$`Number of Suppliers` <- as.character(increment.rows.b$`Number of Suppliers`)
#kable(increment.rows.b)
```

\newpage

## Who were the Top Suppliers in `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`?

```{r, echo=FALSE, fig.height=6}
supplier.agg.py <- filter(supplier.agg, Year == py)
sup.merge.py <- supplier.agg.py[,c(2,4)]
colnames(sup.merge.py) <- c("Supplier", "Prior Yr")
sup.merge <- merge(supplier.agg.cy, sup.merge.py, by.x="Supplier", all.x=TRUE, all.y=FALSE)
sup.merge <- mutate(sup.merge, "Growth" = Total.Sales - `Prior Yr`, "Perc.Growth" = 
                      paste0(round((Growth/`Prior Yr`)*100,1),"%")) %>%
  arrange(desc(Total.Sales)) %>%
  select(1,4:8)
colnames(sup.merge)[c(2,4)] <- c("Sales Current Yr", "Sales Prior Yr")
sup.merge[is.na(sup.merge)] <- 0

### sup.yy here
sup.top10.cy.vec <- unique(sup.merge$Supplier)[1:10]
supplier.agg.10 <- supplier.agg[supplier.agg$Supplier %in% unique(sup.merge$Supplier)[1:10],]
#print(sup.top10.cy.vec)

#horizontal bar chart showing top 10 Suppliers year over year
g <- ggplot(data = supplier.agg.10, aes(x = Supp.By.Year, fill = factor(Year))) + 
  geom_bar(stat = "identity", aes(y = Total.Sales)) + coord_flip() +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_brewer(direction = 1, palette = "RdBu", name = "Year") + theme_dark() + 
  labs(title = "Sales From Top 10 Suppliers By Year", y = "Total Sales", x = "Suppler By Year") 
print(g)

supplier.agg.cy.pretty <- sup.merge
supplier.agg.cy.pretty$Perc.Whole <- paste0(supplier.agg.cy.pretty$Perc.Whole, "%")
supplier.agg.cy.pretty$Growth <- paste0("$",prettyNum(round(supplier.agg.cy.pretty$Growth), big.mark = ","))
supplier.agg.cy.pretty$`Sales Prior Yr` <- paste0("$",prettyNum(round(supplier.agg.cy.pretty$`Sales Prior Yr`), big.mark = ","))
supplier.agg.cy.pretty$`Sales Current Yr` <- paste0("$",prettyNum(round(supplier.agg.cy.pretty$`Sales Current Yr`), big.mark = ",")) ## Switch back
if(nrow(supplier.agg.cy.pretty) < 25){
  kable.suppliers <- supplier.agg.cy.pretty
} else {
  kable.suppliers <- supplier.agg.cy.pretty[1:25,]
}
kable(kable.suppliers)
```


```{r, echo=FALSE}
## Who are the SMALLEST 25% of Suppliers?
#sup.merge <- arrange(sup.merge, `Sales Current Yr`)
#supplier.agg.cy.pretty.b <-  sup.merge   #supplier.agg.cy.bottom
#supplier.agg.cy.pretty.b$Perc.Whole <- paste0(supplier.agg.cy.pretty.b$Perc.Whole, "%")
#supplier.agg.cy.pretty.b$`Sales Current Yr` <-paste0("$",prettyNum(round(supplier.agg.cy.pretty.b$`Sales Current Yr`), big.mark = ","))
#supplier.agg.cy.pretty.b$`Sales Prior Yr` <-paste0("$",prettyNum(round(supplier.agg.cy.pretty.b$`Sales Prior Yr`), big.mark = ","))
#supplier.agg.cy.pretty.b$Growth <-paste0("$",prettyNum(round(supplier.agg.cy.pretty.b$Growth), big.mark = ","))
#kable(supplier.agg.cy.pretty.b[1:109,])

####
# for inline in next title
top.sup1 <- supplier.agg.cy$Supplier[1]
top.sup2 <- supplier.agg.cy$Supplier[2]
top.sup3 <- supplier.agg.cy$Supplier[3]
```

\newpage

## Chart of Monthly Sales for `r top.sup1` - #1 Supplier in `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`
```{r, sup1, echo=FALSE, message=FALSE, warning=FALSE}

sup1.df <- data.clean[data.clean$Supplier == supplier.agg.cy$Supplier[1],]

sup1.ts <- exploratory.jh.time.series(sup1.df, freq = 12)

autoplot(sup1.ts) + theme_dark() + geom_line(color = "#67a9cf", size = 4)+ labs(y = "Sales") +
  scale_y_continuous(labels = scales::dollar)

exploratory.top.three.prod.ts <- function(supplier.filter, df.to.use){
  
  require(dplyr)
  require(fpp2)
  exploratory.jh.time.series <- function(clean.df, freq = 365){
    require(lubridate)
    require(dplyr)
    clean.df <- arrange(clean.df, Date.Sold)
    clean.df$Day <- date(clean.df$Date.Sold) # df should have been arranged in cleaning step
    start.year <- year(clean.df$Day)[1]
    if(freq == 365){
      start.day <- date(head(clean.df$Day,1)) - floor_date(head(clean.df$Day,1), unit = "year")+1 # subtracts earliest day from first day of year plus one
      sales.agg <- aggregate(Total.Sales ~ Day, clean.df, sum)
      all.dates <- seq.Date(from = as.Date(min(sales.agg$Day, na.rm = TRUE)), to = as.Date(max(sales.agg$Day, na.rm = TRUE)),by = 1)
      sales.agg.all <- left_join(data.frame(Day = all.dates), sales.agg)
      sales.ts <- ts(sales.agg.all$Total.Sales, start = c(start.year, start.day), frequency = 365)
    }
    if(freq == 12){
      start.month <- month(head(clean.df$Month,1))
      sales.agg <- aggregate(Total.Sales ~ Month + Year, clean.df, sum)
      #print(head(sales.agg, 25))
      sales.agg.all <- left_join(sales.agg, data.frame(Month = 1:12))
      sales.ts <- ts(sales.agg.all$Total.Sales, start = c(start.year, start.month), frequency = 12)
    }
    sales.ts[is.na(sales.ts)] <- 0
    
    sales.ts
  }
  
  df.one.supplier <- filter(df.to.use, Supplier == supplier.filter)
  agg.by.prod <- aggregate(Total.Sales ~ Description, df.one.supplier, sum)
  agg.by.prod <- arrange(agg.by.prod, desc(Total.Sales))
  top.prods <- agg.by.prod[c(1,2,3), 1]
  
  prod1 <- filter(df.one.supplier, Description == top.prods[1])
  prod1.ts <- exploratory.jh.time.series(prod1, freq = 12)
  #print(head(prod1.ts))
  
  prod2 <- filter(df.one.supplier, Description == top.prods[2])
  prod2.ts <- exploratory.jh.time.series(prod2, freq = 12)
  
  prod3 <- filter(df.one.supplier, Description == top.prods[3])
  prod3.ts <- exploratory.jh.time.series(prod3, freq = 12)
  
  joint.ts <- ts.union(prod1.ts, prod2.ts, prod3.ts)
  colnames(joint.ts) <- top.prods
  #joint.length <- nrow(joint.ts)
  #print(joint.length)
  #label.vec <- seq(from = "1", to = joint.length, by = 3)
  #print(label.vec)
  
  #create labels and breaks
  df.to.use <- arrange(df.to.use, Date.Sold)
  df.to.use <- df.to.use[!is.na(df.to.use$Date.Sold),]
  df.to.use$Day <- date(df.to.use$Date.Sold)
  start.year <- year(df.to.use$Day)[1]
  end.year <- year(df.to.use$Day[length(df.to.use$Day)])

  autoplot(joint.ts) +
theme_dark() + theme(legend.position = "top") + geom_line(size = 2) + scale_color_brewer(direction = -1) + 
    scale_x_continuous(breaks = seq(from = start.year, to = end.year, by = 1)) + 
    labs(y = "Sales", title = paste0("Monthly Sales for the Top 3 Products from ", supplier.filter)) +
           scale_y_continuous(labels = scales::dollar)
  #joint.df <- data.frame(Y=as.matrix(joint.ts), date=time(joint.ts))
  #ggplot(data = joint.df, aes(x = data))

}

exploratory.top.three.prod.ts(top.sup1, data.clean)

```

## Chart of Monthly Sales for `r top.sup2` - #2 Supplier in `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`
```{r, sup2, echo=FALSE, message=FALSE, warning=FALSE}

sup2.df <- data.clean[data.clean$Supplier == supplier.agg.cy$Supplier[2],]

sup2.ts <- exploratory.jh.time.series(sup2.df, freq = 12)

autoplot(sup2.ts) + theme_dark() + geom_line(color = "#67a9cf", size = 4) + labs(y = "Sales") +
  scale_y_continuous(labels = scales::dollar)

exploratory.top.three.prod.ts(top.sup2, data.clean)
```

## Chart of Monthly Sales for `r top.sup3` - #3 Supplier in `r year(as.Date(max(data.clean$Date.Sold, na.rm = TRUE)))`
```{r, sup3, echo=FALSE, message=FALSE, warning=FALSE}

sup3.df <- data.clean[data.clean$Supplier == supplier.agg.cy$Supplier[3],]

sup3.ts <- exploratory.jh.time.series(sup3.df, freq = 12)

autoplot(sup3.ts) + theme_dark() + geom_line(color = "#67a9cf", size = 4) + labs(y = "Sales") +
  scale_y_continuous(labels = scales::dollar)

exploratory.top.three.prod.ts(top.sup3, data.clean)
